Turbulence and Particle Acceleration in Giant Radio Haloes: the Origin of Seed Electrons

### 2016-08-03 paper.tex
Add files via upload

\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\newcommand\cp[1]{{\color{red} #1}}
\newcommand\po[1]{{\color{blue} #1}}

@@

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in cosmological simulations of cluster formation, and confront
  them with radio surface brightness and spectral observations of Coma. For
  standard assumptions, structure formation shocks lead to a seed electron
  population which produces too centrally concentrated radio emission. By
  modifying properties of the CR population (rapid streaming; enhanced CR
  electron acceleration at shocks) or turbulence (increasing
  turbulent-to-thermal energy density with radius), we can match observations,
  but only at the expense of fine-tuning. We then perform a parameter
  study. Under naive assumptions, radio properties are exponentially sensitive
  to the amplitude of turbulence, which is inconsistent with small scatter in
  scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. Understanding the small scatter in radio halo scaling relations
  may prove a rich source of insight on plasma processes in the ICM.
\end{abstract}

@@

While analytic estimates have been made, there has been no ab initio
demonstration that structure formation can lead to the required
abundance of seed electrons with the correct spatial and spectral
characteristics. This is a non-trivial requirement: Coulomb cooling in
dense cluster cores is severe, and DSA fossil electrons may not
survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
Figure~\ref{fig:Edens} we show that such a distribution is not predicted
by cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}. If CRps are predominantly
advected with the cluster plasma, their distribution will be peaked
towards the cluster centre and shows a similar characteristics as the
thermal plasma. As a consequence, the distribution of secondary
electrons and the resulting radio synchrotron emission is also peaked
since the hadronic reaction is a two-body scattering process. Hence,
the simulated emission falls short of the observed extended and flat
radio profile of the Coma cluster.

@@

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. We are not very sensitive to this
assumption, since the thermal and CR quantities are very similar a few 100 Myrs
before and after $z = 0$, where we have chosen to evaluate the simulations. In
all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
\S\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

Thus far, we have ignored CR transport. However, if CRps stream in the ICM, then
their spatial profile could potentially flatten sufficiently
\citep{ensslin11,wiener13}. This scenario is very attractive: it generates seed
electrons with the right spatial footprint, and by removing CRps from the core,
obeys gamma-ray constraints. Turbulence plays two opposing roles: Alfv{\'e}nic
turbulence damps waves generated by the CR streaming instability
\citep{yan02,farmer04}, thus reducing self-confinement; but compressible fast
modes scatter CRs directly. Turbulent damping is still efficient for highly
subsonic conditions \citep{wiener13}, while we assume compressible fast modes to
only provide effective spatial confinement during the periods of transsonic,
highly super-Alfv{\'e}nic ($\mathcal{M}_{\rm A} \sim 5$) turbulence associated with
mergers. Thus, CRs can stream out when the cluster is kinematically
quiescent. Furthermore, even Alfv{\'e}nic streaming timescales are relatively
short \cite[$\sim 0.1-0.5$ Gyr;][]{wiener13} compared to the timescale on which
the CRp population is built up. Based on these findings, we adopt a toy model
for our \Mstream scenario in which CR streaming instantaneously produces flat
CRp profiles. We assume that CRs cannot stream significantly past perpendicular
$B$-fields at the accretion shock, so that the total number of CRs is conserved
within the virial radius during the streaming process.  In all our calculations
we assume that turbulent reacceleration efficiently accelerates particles for
$\tau_\rmn{cl}\sim650$~Myrs (which is roughly the cascade time on which
turbulence is damped) and that during this turbulent phase CR streaming and
spatial diffusion can be neglected.

@@

The important effects are best summarised in terms of the acceleration \cp{rate
  which is goverend by advection in momentum space:
\begin{equation}
  \tau_{\rm D}^{-1}\equiv \Gamma_{\rm D}\equiv
  \frac{\dot{p}}{p}=p^{-3}\frac{\partial}{\partial p}\left(p^2D_{pp}\right)
  =\frac{4 D_{pp}}{p^{2}}.
\end{equation}
In the last step we have used that $D_{pp}\propto p^2$ for turbulent
reacceleration (equation \ref{eqn:Dpp}). Hence the acceleration time is independent of momentum.}
%\begin{equation}
%\tau_{\rm D}= \frac{p^{2}}{4 D_{pp}}
%\end{equation}
%which is momentum independent.
This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use $B$-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_L=2\upi/\lambda_L$, where $\lambda_L = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This length scale corresponds to an eddy turnover time on the outer scale of \cp{$2\upi\lambda_L\varv_L^{-1} \sim 1.2$~Gyr} if $\varv_L \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_L$, in some cases comparable to the size of the cluster \citep{vazza11,miniati15}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_L$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parametrized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_{\rmn{e}}c)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_{\rmn{p}}c)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_{\rmn{e}}c)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_{\rmn{e}}/m_{\rmn{p}}$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We use a simple test-particle model for the CR acceleration and
injection, where each shock injects CRs that trace a power-law in
momentum,
\begin{equation}
  f_\p(p,t) = C(t)\,p^{\alpha_\rmn{inj}}\,,\quad
  \alpha_\rmn{inj}=\frac{(\gamma_\rmn{ad}+1)\mathcal{M}^2}
        {(\gamma_\rmn{ad}-1)\mathcal{M}^2+2}
\end{equation}
determined by the normalisation $C(t)$ and the spectral index
$\alpha_{\rmn{inj}}$ that depends on the adiabatic index
$\gamma_\rmn{ad} = 5/3$ and the Mach number of the shock
$\mathcal{M}$. It is given by the ratio of the upstream velocity
($\varv_2$) and the sound speed ($c_{\rm s}$). The CR number density and CR
energy density are derived from
\begin{eqnarray}
n_\CR &=&
\int_{p_\rmn{inj}}^\infty \dd p\, f_\p(p)\\
\eps_\CR &=&
\int_{p_\rmn{inj}}^\infty \dd p\, f_\p(p) \,E(p),
\end{eqnarray}
where $E(p) = (\sqrt{1+p^2} -1)\, m_{\rmn{p}}\,c^2$ is the kinetic energy of a
proton with momentum $p$. We adopt a fit to Monte Carlo simulations of the
thermal leakage process that relates the momentum of injected protons
($p_\rmn{inj}$) to the thermal energy ($p_\rmn{th}$) of the shocked plasma
\citep{kang11}:
\begin{eqnarray}
  \label{eq:qinj}
  p_\rmn{inj} &=& x_\rmn{inj} p_\rmn{th} =
  x_\rmn{inj} \sqrt{\frac{2 \,\kB T_2}{m_{\rmn{p}} c^2}}\,, \nonumber \\
  \rmn{where}\quad x_\rmn{inj} &\approx& 1.17 \frac{\varv_2}{p_{\rmn{th}}\,c} \left(1+
  \frac{1.07}{\eb}\right) \left(\frac{\mathcal{M}}{3}\right)^{0.1}\,.
\end{eqnarray}
Here $\eb = B_0/B_{\perp}$, $B_0$ is the amplitude of the downstream
MHD wave turbulence, and $B_{\perp}$ is the magnetic field along the
shock normal. The physical range of $\eb$ is quite uncertain due to
complex plasma interactions. In this paper, we adopt $\eb = 0.23$,
which -- as we will later see -- corresponds to a conservative maximum
energy acceleration efficiency for protons of $0.1$. To derive the
acceleration efficiency, $\zeta_\rmn{inj}$, we first have to infer the
particle injection efficiency, which is the fraction of downstream
thermal gas particles which experience diffusive shock acceleration
\citep[for details see][]{pinzke13},
\begin{equation}
  \label{eq:eta}
  \eta_\rmn{p,lin} =
  \frac{4}{\sqrt{\upi}}\,\frac{x_\rmn{inj}^3}{\alpha_\rmn{inj}-1}\,
  \rmn{e}^{-x_\rmn{inj}^2}.
\end{equation}
The particle injection efficiency is a strong function of
$x_\rmn{inj}$ that depends on both $\mathcal{M}$ and $\eb$. The
energy density of CRs that are injected and accelerated at the shock
(neglecting the CR back reaction on the shock) is given by
\begin{equation}
\label{eq:CR_energy} 
  \Delta\eps_\rmn{CR,lin} =
  \eta_\rmn{p,lin}(\mathcal{M})\,n_\rmn{th}(T_2)
  \,\frac{\eps_\CR}{n_\CR}
\end{equation}
and the CR energy injection and acceleration efficiency is:
\begin{equation}
  \zeta_\rmn{lin} =
  \frac{\Delta\eps_\rmn{p,lin}}{\Delta\eps_\rmn{diss}},
   \quad\mbox{where}\quad
  \Delta\eps_\rmn{diss} = \eps_\rmn{th2} - \eps_\rmn{th0}\,\left(\frac{\rho_2}{\rho_0}\right)^{\gamma_\rmn{ad}}\,.
\label{eqn:energy_frac}  
\end{equation}
The dissipated energy density in the downstream regime,
$\Delta\eps_\rmn{diss}$, is given by the difference of the thermal
energy densities in the pre- and post-shock regimes, corrected for the
adiabatic energy increase due to gas compression.

@@

In Figure~\ref{fig:sync_profile}, we show radial profiles for the radio
emission in all three scenarios in which the seeds undergo Fermi-II
reacceleration in turbulent fields that are shaped such as to
reproduce the Coma RH profile at 352~MHz.  Adopting our
parametrization for the volumetric injection rate of turbulent energy,
$I_0\propto \eps_\rmn{th}^{\alpha_\rmn{tu}}$, we find that to fit the observations, we require
$\alpha_\rmn{tu}= 0.67$ for \Mflatturb, $\alpha_\rmn{tu}= 0.82$ for
\Mstream, and $\alpha_\rmn{tu}= 0.88$ for \Mprimary. As a result, the
ratio of turbulent-to-thermal energy densities slightly increase with
radius as shown in Figure~\ref{fig:turb}.  The fine-tuning of these exponents is somewhat problematic, as we discuss in \S\ref{sect:param_comp}. After turbulent
reacceleration, the volume-weighted, relative CRp energy density and
relative CRp number density inside the RH for \Mflatturb (\Mstream),
are found to be 2 (3) per cent and $2\times10^{-8}$ ($5\times10^{-8}$),
respectively.

@@

Figure~\ref{fig:sync_profile} demonstrates that the modelled radio
profiles without turbulent reacceleration are too steep.  In the
bottom right panel of Figure~\ref{fig:sync_profile} (labelled with
Brunetti et al. 2012), we show that our simulated profiles of
reaccelerated CRs, which only take advective CR transport into
account, i.e. they neglect CR streaming or a flatter turbulent
profile, produce radio profiles that are too steep. Indeed, even using
the assumptions of previous work -- where complete freedom in the seed
population was allowed -- it is not possible to reproduce observations
in both frequencies in any model.\footnote{Note that in previous work
  on the Coma cluster, $\eps_\rmn{turb} \propto \eps_\rmn{th}$ was
  adopted which approximately corresponds to $\alpha_\rmn{tu}= 1$
  \citep{brunetti12} and together with the different distributions for
  seed CRes constitute the main differences compared to our work.}
Decreasing the acceleration efficiency with radius does not change
this conclusion much because of the weak radial dependence of
$D_{pp}(R)\propto \eps_\rmn{th}(R)^{\alpha_\rmn{tu}-1}
\sqrt{T(R)}$. This signals that the problem is generic and requires
either additional modifications to the plasma physics of acceleration
or a better understanding of potential observational systematics. In
addition there are differences in the simulated density and
temperature profiles in comparison to the observed profile in Coma
that impact the CR abundance as well as cooling and reacceleration.

>>>>

\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\newcommand\cp[1]{{\color{red} #1}}
\newcommand\po[1]{{\color{blue} #1}}
\newcommand\AP[1]{{\color{green} #1}}

@@

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in \AP{the post-processing of} cosmological simulations of cluster formation, and confront
  them with radio surface brightness and spectral observations of Coma. For
  standard assumptions, structure formation shocks lead to a seed electron
  population which produces too centrally concentrated radio emission. By
  modifying properties of the CR population (rapid streaming; enhanced CR
  electron acceleration at shocks) or turbulence (increasing
  turbulent-to-thermal energy density with radius), we can match observations,
  but only at the expense of fine-tuning. We then perform a parameter
  study. Under naive assumptions, radio properties are exponentially sensitive
  to the amplitude of turbulence, which is inconsistent with small scatter in
  scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. Understanding the small scatter in radio halo scaling relations
  may prove a rich source of insight on plasma processes in the ICM.
\end{abstract}

@@

While analytic estimates have been made, there has been no ab initio
demonstration that structure formation can lead to the required
abundance of seed electrons with the correct spatial and spectral
characteristics. This is a non-trivial requirement: Coulomb cooling in
dense cluster cores is severe, and DSA fossil electrons may not
survive. On the other hand, for secondaries to constitute the seed
population, the CRp population required in the best-studied case of
the Coma cluster must have a very broad and flat (or even slightly
inverted) spatial profile \citep{brunetti12}, in contrast with the
thermal plasma whose energy density declines steeply with radius. In
Figure~\ref{fig:Edens} we show that such a distribution is not predicted
by cosmological simulations \cite[see
  also][]{pinzke10,2014MNRAS.439.2662V}. If CRps are predominantly
advected with the cluster plasma, their distribution will be peaked
towards the cluster centre and show a similar characteristics as the
thermal plasma. As a consequence, the distribution of secondary
electrons and the resulting radio synchrotron emission is also peaked
since the hadronic reaction is a two-body scattering process. Hence,
the simulated emission falls short of the observed extended and flat
radio profile of the Coma cluster.

@@

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. \AP{Note that in recent simulations by \citet{miniati15}, the turbulent reacceleration is strongest around core passage. However, we are not very sensitive to the adopted decay time, since the thermal and CR quantities are very similar a few 100 Myrs before and after $z = 0$, where we have chosen to evaluate the simulations.}
In all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
\S\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

Thus far, we have ignored CR transport. However, if CRps stream in the ICM, then
their spatial profile could potentially flatten sufficiently
\citep{ensslin11,wiener13}. This scenario is very attractive: it generates seed
electrons with the right spatial footprint, and by removing CRps from the core,
obeys gamma-ray constraints. Turbulence plays two opposing roles: Alfv{\'e}nic
turbulence damps waves generated by the CR streaming instability
\citep{yan02,farmer04}, thus reducing self-confinement; but compressible fast
modes scatter CRs directly. Turbulent damping is still efficient for highly
subsonic conditions \citep{wiener13}, while we assume compressible fast modes to
only provide effective spatial confinement during the periods of transsonic,
highly super-Alfv{\'e}nic ($\mathcal{M}_{\rm A} \sim 5$) turbulence associated with
mergers. Thus, CRs can stream out when the cluster is kinematically
quiescent. Furthermore, even Alfv{\'e}nic streaming timescales are relatively
short \cite[$\sim 0.1-0.5$ Gyr;][]{wiener13} compared to the timescale on which
the CRp population is built up. Based on these findings, we adopt a toy model
for our \Mstream scenario in which CR streaming instantaneously produces flat
CRp profiles. We assume that CRs cannot stream significantly past perpendicular
$B$-fields at the accretion shock, so that the total number of CRs is conserved
within the virial radius during the streaming process.

@@

The important effects are best summarised in terms of the acceleration \cp{rate
  which is goverend by advection in momentum space:
\begin{equation}
  \tau_{\rm D}^{-1}\equiv \Gamma_{\rm D}\equiv
  \frac{\dot{p}}{p}=p^{-3}\frac{\partial}{\partial p}\left(p^2D_{pp}\right)
  =\frac{4 D_{pp}}{p^{2}}.
\end{equation}
In the last step we have used that $D_{pp}\propto p^2$ for turbulent
reacceleration (equation \ref{eqn:Dpp}). Hence the acceleration time is independent of momentum.}
%\begin{equation}
%\tau_{\rm D}= \frac{p^{2}}{4 D_{pp}}
%\end{equation}
%which is momentum independent.
This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use $B$-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_L=2\upi/\lambda_L$, where $\lambda_L = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This length scale corresponds to an eddy turnover time on the outer scale of \cp{$2\upi\lambda_L\varv_L^{-1} \sim 1.2$~Gyr} if $\varv_L \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_L$, in some cases comparable to the size of the cluster \AP{\citep[e.g. $\lambda_L \sim 1 {\rm Mpc}$ in ][]{vazza11,miniati15}}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_L$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parametrized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_{\rmn{e}}c)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_{\rmn{p}}c)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_{\rmn{e}}c)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_{\rmn{e}}/m_{\rmn{p}}$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

@@

We use a simple test-particle model for the CR acceleration and
injection, where each shock injects CRs that trace a power-law in
momentum,
\begin{equation}
  f_\p(p,t) = C(t)\,p^{\alpha_\rmn{inj}}\,,\quad
  \alpha_\rmn{inj}=\frac{(\gamma_\rmn{ad}+1)\mathcal{M}^2}
        {(\gamma_\rmn{ad}-1)\mathcal{M}^2+2}
\end{equation}
determined by the normalisation $C(t)$ and the spectral index
$\alpha_{\rmn{inj}}$ that depends on the adiabatic index
$\gamma_\rmn{ad} = 5/3$ and the Mach number of the shock
$\mathcal{M}$ \AP{\citep[see also ][]{1998ApJ...502..518Q,miniati01,pfrommer06}}. It is given by the ratio of the upstream velocity
($\varv_2$) and the sound speed ($c_{\rm s}$). The CR number density and CR
energy density are derived from
\begin{eqnarray}
n_\CR &=&
\int_{p_\rmn{inj}}^\infty \dd p\, f_\p(p)\\
\eps_\CR &=&
\int_{p_\rmn{inj}}^\infty \dd p\, f_\p(p) \,E(p),
\end{eqnarray}
where $E(p) = (\sqrt{1+p^2} -1)\, m_{\rmn{p}}\,c^2$ is the kinetic energy of a
proton with momentum $p$. We adopt a fit to Monte Carlo simulations of the
thermal leakage process that relates the momentum of injected protons
($p_\rmn{inj}$) to the thermal energy ($p_\rmn{th}$) of the shocked plasma
\citep{kang11}:
\begin{eqnarray}
  \label{eq:qinj}
  p_\rmn{inj} &=& x_\rmn{inj} p_\rmn{th} =
  x_\rmn{inj} \sqrt{\frac{2 \,\kB T_2}{m_{\rmn{p}} c^2}}\,, \nonumber \\
  \rmn{where}\quad x_\rmn{inj} &\approx& 1.17 \frac{\varv_2}{p_{\rmn{th}}\,c} \left(1+
  \frac{1.07}{\eb}\right) \left(\frac{\mathcal{M}}{3}\right)^{0.1}\,.
\end{eqnarray}
Here $\eb = B_0/B_{\perp}$, $B_0$ is the amplitude of the downstream
MHD wave turbulence, and $B_{\perp}$ is the magnetic field along the
shock normal. The physical range of $\eb$ is quite uncertain due to
complex plasma interactions. In this paper, we adopt $\eb = 0.23$,
which -- as we will later see -- corresponds to a conservative maximum
energy acceleration efficiency for protons of $0.1$. To derive the
acceleration efficiency, $\zeta_\rmn{inj}$, we first have to infer the
particle injection efficiency, which is the fraction of downstream
thermal gas particles which experience diffusive shock acceleration
\citep[for details see][]{pinzke13},
\begin{equation}
  \label{eq:eta}
  \eta_\rmn{p,lin} =
  \frac{4}{\sqrt{\upi}}\,\frac{x_\rmn{inj}^3}{\alpha_\rmn{inj}-1}\,
  \rmn{e}^{-x_\rmn{inj}^2}.
\end{equation}
The particle injection efficiency is a strong function of
$x_\rmn{inj}$ that depends on both $\mathcal{M}$ and $\eb$. The
energy density of CRs that are injected and accelerated at the shock
(neglecting the CR back reaction on the shock) is given by
\begin{equation}
\label{eq:CR_energy} 
  \Delta\eps_\rmn{CR,lin} =
  \eta_\rmn{p,lin}(\mathcal{M})\,n_\rmn{th}(T_2)
  \,\frac{\eps_\CR}{n_\CR}
\end{equation}
and the CR energy injection and acceleration efficiency is:
\begin{equation}
  \zeta_\rmn{lin} =
  \frac{\Delta\eps_\rmn{p,lin}}{\Delta\eps_\rmn{diss}},
   \quad\mbox{where}\quad
  \Delta\eps_\rmn{diss} = \eps_\rmn{th2} - \eps_\rmn{th0}\,\left(\frac{\rho_2}{\rho_0}\right)^{\gamma_\rmn{ad}}\,.
\label{eqn:energy_frac}  
\end{equation}
The dissipated energy density in the downstream regime,
$\Delta\eps_\rmn{diss}$, is given by the difference of the thermal
energy densities in the pre- and post-shock regimes, corrected for the
adiabatic energy increase due to gas compression.

@@

In Figure~\ref{fig:sync_profile}, we show radial profiles for the radio
emission in all three scenarios in which the seeds undergo Fermi-II
reacceleration in turbulent fields that are shaped such as to
reproduce the Coma RH profile at 352~MHz.  Adopting our
parametrization for the volumetric injection rate of turbulent energy,
$I_L\propto \eps_\rmn{th}^{\alpha_\rmn{tu}}$, we find that to fit the observations, we require
$\alpha_\rmn{tu}= 0.67$ for \Mflatturb, $\alpha_\rmn{tu}= 0.82$ for
\Mstream, and $\alpha_\rmn{tu}= 0.88$ for \Mprimary. As a result, the
ratio of turbulent-to-thermal energy densities slightly increase with
radius as shown in Figure~\ref{fig:turb}.  The fine-tuning of these exponents is somewhat problematic, as we discuss in \S\ref{sect:param_comp}. After turbulent
reacceleration, the volume-weighted, relative CRp energy density and
relative CRp number density inside the RH for \Mflatturb (\Mstream),
are found to be 2 (3) per cent and $2\times10^{-8}$ ($5\times10^{-8}$),
respectively.

@@

Figure~\ref{fig:sync_profile} demonstrates that the modelled radio
profiles without turbulent reacceleration are too steep.  In the
bottom right panel of Figure~\ref{fig:sync_profile} (labelled with
Brunetti et al. 2012), we show that our simulated profiles of
reaccelerated CRs, which only take advective CR transport into
account, i.e. they neglect CR streaming or a flatter turbulent
profile, produce radio profiles that are too steep. \AP{Indeed, even using
the assumptions of previous work, where complete freedom in the seed
population was allowed, it is not possible to reproduce observations
in both frequencies in any model\footnote{Note that in previous work
  on the Coma cluster, $\eps_\rmn{turb} \propto \eps_\rmn{th}$ was
  adopted which approximately corresponds to $\alpha_\rmn{tu}= 1$
  \citep{brunetti12} and together with the different distributions for
  seed CRes constitute the main differences compared to our work.} -- with or without turbulent reacceleration.}
Decreasing the acceleration efficiency with radius does not change
this conclusion much because of the weak radial dependence of
$D_{pp}(R)\propto \eps_\rmn{th}(R)^{\alpha_\rmn{tu}-1}
\sqrt{T(R)}$. This signals that the problem is generic and requires
either additional modifications to the plasma physics of acceleration
or a better understanding of potential observational systematics. In
addition there are differences in the simulated density and
temperature profiles in comparison to the observed profile in Coma
that impact the CR abundance as well as cooling and reacceleration.
